---
title: "StoX FDA data preparation (baseline)"
author: "Edvin Fuglebakk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{StoX FDA data preparation (baseline)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE
)
library(RstoxFDA)
```

This vignette introduces common tasks in data preparation for fisheries dependent analysis (FDA), and the functions that RstoxFDA provides for "baseline"-models in the StoX framework to support these tasks. Familiarity with the StoX framework is assumed. These functions are mostly motivated by needs in fisheries-dependent analysis at the Institute of Marine Research (IMR), and supports common data-transformation, annotation and code translation. FDA at IMR typically involve using census data from the Norwegian Directorate of Fisheries (FDIR) and reporting to the International Counsel for Exploration of the Seas (ICES). The data-preparation functions will be introduced in the context of setting up the StoX baseline for catch at age estimation with the package 'Reca', but most tasks are relevant for other estimation methods or reporting as well. 

This document introduces some problems and tasks, and how RstoxFDA-functions may be applied to solve them. The details of how to use the functions introduced here is provided as function documentation viewable in the StoX "function description"-tab, or in an R-console via '?', e.g:

```{r docfunction, echo=TRUE}
?RstoxFDA::ConvertWeightBiotic
```

Documentation for data formats are provided in the same way. E.g.:

```{r docformat, echo=TRUE}
?RstoxBase::StratumPolygon
```


As StoX is composed of several R-packages. Functions will be referred to by their package name with the notation _package::function_. In the StoX user interface, the package that functions belong to are not visible. Any data formats are similarly denoted **package::format**.

## Obtaining data
While this vignette introduces some tools to assist correct usage of FDA-data, no attempt is made at introducing unfamiliar users to the peculiarities of fisheries dependent sampling. For IMR samples a provisional guide to FDA-data (in Norwegian) can be found at the [IMR intranet: Kilder fiskeriavhengige data](https://havforskningsinstituttet.sharepoint.com/sites/Fiskeridynamikk/Delte%20dokumenter/Forms/AllItems.aspx?id=%2Fsites%2FFiskeridynamikk%2FDelte%20dokumenter%2FKilder%20%E2%80%93%20fiskeriavhengige%20data%2Epdf&parent=%2Fsites%2FFiskeridynamikk%2FDelte%20dokumenter), and for census data some introduction (in Norwegian) can be found at the [IMR intranet: Fiskeridata](https://havforskningsinstituttet.sharepoint.com/sites/Forskningsdata/SitePages/Fiskeridata.aspx)

Within the IMR-computer-network both landing data and sample data can be obtained from the [dataset explorer](https://datasetexplorer.hi.no/apps/datasetexplorer/v2/navigation). Consider using the tools-menu to download landing data and to compile sample data, if the pre-compiled data sets under 'commercial sampling' are not suitable. Otherwise landing data is available in various formats as described at the [IMR intranet: Fiskeridata](https://havforskningsinstituttet.sharepoint.com/sites/Forskningsdata/SitePages/Fiskeridata.aspx), and some data is publicly available directly from [FDIR](https://www.fiskeridir.no/Tall-og-analyse/AApne-data) (also in Norwegian).

## Installation
RstoxFDA is an optional package in StoX. In order to use the functions introduced in this vignette in the StoX user interface, make sure RstoxFDA is installed. See installation instructions on [github: StoXProject/RstoxFDA](https://github.com/StoXProject/RstoxFDA). To install the StoX user interface, see instructions on [github: StoXProject/StoX](https://github.com/StoXProject/StoX)

# Data formats in StoX
StoX support several types of input, such as records of biological measurements (biotic-data), records of acoustic observations (acoustic-data) and records of total weight landed (landing-data). For FDA, it is mainly biotic-data and landing-data, that are of interest. For each of these types of input, StoX may support several file-formats, but most of the data-processing and visualization supported by StoX requires that the various input format are converted to StoX-specific standards. Biotic-data must be converted to **RstoxData::StoxBioticData**, which can be done by the function _RstoxData::StoxBiotic_, and landing data must be converted to **RstoxData::StoxLandingData**, which is can be done the function _RstoxData::StoxLanding_. This makes sure that StoX-functions and visualisations in the StoX user-interface can rely on some unified definitions for important variables. In cases when a unified definition is not needed, StoX functions may be configured to handle other columns than the standard variables. Such columns may then be added to the standard StoX-formats by use of functions like _RstoxData::AddToStoxBiotic_.

In addition to the data types, parameters and configurations may be read from various kind of 'resource-files'. The format of these files are described in the documentation for the functions that import them. E.g:

```{r resourceformat, echo=TRUE}
?RstoxFDA::DefineLengthConversionParameters
```


# Sample data
StoX supports several input formats for biological samples, which can be read by the function _RstoxData::ReadBiotic_. This function reads biotic data into an internal structure, **RstoxData::BioticData**, which is rather loosely defined. Data are organized into several tables with different columns, and include some information about the source of the data, but otherwise **RstoxData::BioticData** does not provide a very strict format. Biotic data therefore have to be converted to **RstoxData::StoxBioticData**, and it is usually desirable to convert your biological data to that format with the function _RstoxData::StoxBiotic_. Some data transformation, filtering, and re-coding is necessary to do before converting to **RstoxData::StoxBioticData**. Since the **RstoxData::BioticData** is defined in a very flexible manner, some of the functions that are provided to perform such data-transformations and re-codings only make sense for particular sources of the biotic-data. Since the source of biotic-data at IMR typically is NMDbiotic, some functions specific to handling that data source will be introduced here.

### Filtering
Some data-preparation or filtering is necessary to do before converting to **RstoxData::StoxBioticData**, and some are necessary to do after converting. Data-preparation that can be done both before and after converting is generally desirable to do before, since error messages then provide a clearer link to the raw data, and since it generally reduces the amount of data that needs to be further processed. However, there are also some concerns to consider with respect-to re-use of StoX-projects. StoX is typically used for analyses that are performed on an annual basis, with minor modifications to the data preparation for each year. In this setting it is useful to be able to make a copy of a StoX-project from a previous year, and then make necessary modifications to that copy. It should then be kept in mind, that **RstoxData::BioticData** contains tables with names corresponding to input data-files, and input data-files should ideally have different names for different years. This means that filters configured by _RstoxData::FilterBiotic_, must typically all be changed from one year to the next, while filters configured with _RstoxData::FilterStoxBiotic_ are typically reusable.

## NMDbiotic
Biological samples at IMR are archived in a versioned data-format called NMDbiotic (https://www.imr.no/formats/nmdbiotic/). If the source for your biological data is NMDbiotic, you may find that information is recorded somewhat differently than what allows for straightforward conversion with _RstoxData::StoxBiotic_. For instance, you may find that fishing date is recorded as the end of the fishing operation, that not all catches have a position, or that fish weights are measured of gutted fish, rather than unprocessed fish. And you will find that gear codes are different from those used in the landing-data. RstoxData provide some functions for dealing with these kind of issues, which are presented below. These functions are specific to **RstoxData::BioticData** that are read from NMDbiotic files. If the source if something else (e.g. ICES biotic) it will usually not make sense to apply these functions, and the exact behaviour is specified in the function documentation.

If not all data in the input file are relevant for analysis, the number of data issues to handle can be greatly reduced by first applying the necessary filtering with _RstoxData::FilterBiotic_. As mentioned above, use of _RstoxData::FilterBiotic_ rather than _RstoxData::FilterStoxBiotic_, does come with the caveat that filters are less re-usable if the same analysis is to be done for several years of data. So a good rule of thumb is to try to solve standard filters, that will apply for coming years as well in _RstoxData::FilterStoxBiotic_, possibly through the use of _RstoxData::AddToStoxBiotic_.

### Fishing time information
you may find that fishing date is recorded as the end of the fishing operation, since that was the day the catch was sampled, and that the corresponding fields for the start of the fishing operation is missing. You may also find that the time (hour and minute) is not filled, since that detail may not be available to samplers. Unfortunately, both of these fields ('stationstartdate' and 'stationstarttime' in **RstoxData::BioticData** / NMDbiotic) must be set in order for the conversion with _RstoxData::StoxBiotic_ to preserve time information. The downstream analysis may be as simple as annotating quarter or month to the data, and in that case it is clear that it is not warranted to know the hour and minute of the catch, or to distinguish the start of the fishing operation from the end of it. For situations like these, the function _RstoxFDA::SetStartDateBiotic_ allows you to impute the start date so that it becomes equal to the stop date of the fishing operation, and the function _RstoxFDA::SetTimeBiotic_ allows you to set an arbitrary time of the day to records where time is missing.

### Position information
Some sampling still record position of catches only in terms of area codes. In order for visualisations and annotation of other area codes to be supported by **RstoxData::StoxBioticData**, positional data must be provided. If the area codes provided in the NMDbiotic records provide sufficient precision, you may consider imputing positions from area codes, using _RstoxFDA::SetAreaPositionsBiotic_. See the section 'Area codes' for more details. Before that can be done, some particularities of the IMR data archive may need to be handled. Catch locations are often provided via the 'FDIR Statistical areas and locations', which is a two-code system where the first code represents an area (NMDbiotic: area), and the second a finer area coding system within that area (NMDbiotic: location). Both of these codes are defined as integers, and FDIR uses the convention to prefix single digit integers with a leading '0', E.g: '01' or '09'. In IMR records, several conventions have been used and the area '01' may be recorded as either '01', '1' or ' 1' (note the whitespace), and perhaps additional variants. This may be corrected by using _RstoxBase::DefineTranslation_ and _RstoxBase::TranslateBiotic_.

In addition to this issue, FDIR revised their coding system in 2018, following a similar revision to the ICES areas. In NMDbiotic the old and revised definitions are represented as to different systems of area-coding (signified by the field NMDbiotic: system), but they are nearly identical and are often used interchangeably. _RstoxFDA::SetAreaPositionsBiotic_ may therefore have to occur twice in a StoX project. Once for system '2' (old definition) and once for system '10' (new definition).

### Gear codes
NMDbiotic uses a four digit gear-coding system which is organized so that the two first digits code for gear-groups. If that coarser gear-definition is sufficient for analysis, downstream processing may be simplified by re-coding the gear column to only the first two digits. This is supported with the function _RstoxFDA::SetShortGearBiotic_.

### Haul identification
In order to get position and time attributed to fishing operations in the standard StoX way, it may sometimes be necessary to introduce some identifiers in the field 'station' in **RstoxData::BioticData** / NMDbiotic, before converting to StoX biotic. If fishing stations are known to be identified with fishing operations in a data set, the function _RstoxFDA::CopyBiotic_ may be used to copy the column 'serialnumber' into the column 'station'. This may be considered if you get warnings in StoX baseline that sayse something along the lines of 'There are more than one 'serialnumber' (HaulKey in StoxBioticData) for x out of y 'station'(StationKey in StoxBioticData) in the NMDBiotic data.'. A brief explanation is given below.

Hauls or fishing operations are in NMDbiotic recorded as fishstations and are identified with the serialnumber. The concept of a station, which is a sampling unit, is recorded on the same table and usually there are only one fishing operation for each station, which makes it unnecessary to distinguish. When several hauls are to be recorded for the same station, they are given the same station number. Because this is not common, the use of station numbers have not always been very disciplined by users contributing data to NMDbiotic. This has lead to the field 'station' being left blank, or repeated by mistake. The conversion to **RstoxData::StoxBioticData** is made in accordanse with the standards and since this format has haul and station as different levels, all hauls with the same or missing value in the 'station' field will be attributed to the same station. Moreover, the standard allocation of time and position is in **RstoxData::StoxBioticData** done to station, rather than haul.

Alternatively, time and position may be added to the haul level in **RstoxData::StoxBioticData** via the function _RstoxData::AddToStoxBiotic_.

These workarounds are mainly intended for dealing with historical data. Registration practices should be dieallt be ammended when inconventional use of the field station is encountered.

### Non standard fish measurments
The practicalities of fisheries sampling often dictate that partially processed products, such as gutted fish, are measured. **RstoxData::StoxBioticData** only supports Live (Round) weights and total-length measurements, but weights in FDA-data are often provided as gutted (with or without head) and occasionally only head-lengths are provided (along with age). If conversions via scalar factors or simple regressions are acceptable, data can be made fit for _RstoxData::StoxBiotic_ by use of the functions _RstoxFDA::ConvertWeightBiotic_ and _RstoxFDA::ConvertLengthBiotic_. Conversion tables for these functions must first be defined by the functions _RstoxFDA::DefineLengthConversionParameters_, and _RstoxFDA::DefineWeightConversionFactor_. The latter offers the selection of some built-in official conversion factors, that can be used. Otherwise conversion tables must be provided as resource files. 

### Quality filters
It has been common practice to remove extreme outliers from data. As most of these outliers are erroneous records, it is desirable to remove them from database records, or correct them if information is available, rather than removing them in analysis. RstoxFDA provides some outlier-filters. These are mainly provided in order to reproduce legacy workflows, and may be relevant to apply mostly to older data. They may also be used in conjunction with report functions to verify that extreme outliers are uncommon or absent. The function _RstoxFDA::FilterAgeLengthOutliersStoxBiotic_ provides a filter with a von-Bertalannfy regression for removing extreme age-length relationships. The function _RstoxFDA::FilterWeightLengthOutliersStoxBiotic_ provides a filter with a log-linear model to remove extreme weight-length relationships.

When records are missing, or suspect data are encountered, it is important to provide feedback to data-collectors, so that they can correct the record when possible. In order to progress with analysis, and possibly detecting more data-issues, one may want to filter out offending data, and include them again later if they are corrected in databases. Dedicated filters for this purpose is recommended (_RstoxData::FilterBiotic_). If these filters are organised carefully, a report of what is filtered out may be generated by the function _RstoxFDA::ListBioticDifference_, which lists the difference between the output of a processes that proivde **RstoxData::StoxBioticData** to one providing data in **RstoxData::BioticData**. That kind of workflow allows you to discover several issues before reporting back to data-collectors, and provide a convenient format for reporting issues.

##  Example: StoX-project
In summary, depending on the exact data issues at hand, sample data in a StoX project may be handled by a sequence of processes exemplified in the following table:

| Process name | Function | Description |
|-------------|---------|--------|
| ReadSamples | ReadBiotic | Reads in **RstoxData::BioticData** with biological samples from NMDbiotic xml-files |
| FilterSamples | FilterBiotic | Reduces data set to relevant sampling programs, relevant species, relevant catch fractions, etc. |
| SetTime | SetTimeBiotic | Sets 'stationstarttime' on the table 'fishstation' in **RstoxData::BioticData** to a fixed time of day, wherever 'stationstarttime' is not already set |
| SetStartDate | SetStartDateBiotic | Sets 'stationstartdate' on the table 'fishstation' in **RstoxData::BioticData** equal to 'stationstopdate', wherever 'stationstartdate' is not already set  |
| LoadConversionFactors | DefineWeightConversionFactor | Loads built-in official conversion factors |
| ConvertWeight | ConvertWeightBiotic | Converts weight-fields in **RstoxData::BioticData** to desired product type, using the table defined by LoadConversionFactors |
| ReadLengthParameters | DefineLengthConversionParameters | Reads regression parameters for converting length |
| ConvertLength | ConvertLengthBiotic | Converts 'length' on the table 'individual' in **RstoxData::BioticData** to desired measurement type, using parameters defined by ReadLengthParameters |
| StoxBiotic | StoxBiotic | Converts **RstoxData::BioticData** to **RstoxData::StoxBiotic** |

# Landing data
FDA analysis very often involve some kind of ratio-estimation dependent on a census of landed weight by species, or fishing effort by gear. At IMR we typically get such census data via mandatory reporting to FDIR. While RstoxData provides support for reading sales-notes from FDIR in a standardized format via the function _ReadLanding_, it is often convenient to use other formats. RstoxFDA therefore offers an alternative function for reading sales-notes: _ReadLandingFDA_ that allows for some additional formats, some of these only available internally at IMR. _ReadLandingFDA_ returns data that is compatible with _RstoxData::ReadLanding_ and can be used in any function that accepts **RstoxDAta::LandingData**. While RstoxData offers the function _RstoxData::FilterLanding_, that provides filtering options for that format, the StoX user interface performs best if data is converted to a more aggregated format before further processing. The RstoxData function _RstoxData::StoxLanding_ aggregates sales-notes to the format **RstoxDAta::StoxLandingData** and is supported by most further processing.

##  Example: StoX-project
Typically, landings data are included in the StoX project by a sequence of processes exemplified by the table below:

| Process name | Function | Description |
|-------|------|---------|--------|
| ReadLandings | ReadLandingFDA | Reads in **RstoxData::LandingData** with sales-notes data from one of several available file formats |
| StoxLanding | StoxLanding |Converts **RstoxData::LandingData** to the aggregated **RstoxData::StoxLandingData** |
| FilterLandings | FilterStoxLanding | Reduces data to relevant sampling frame, removing landings by area, vessel flag etc. |


# Area codes
Area codes are commonly used for stratification or domain definitions, and FDA frequently encounter situations where a clear definition of areas are useful. This is provided in StoX as spatial strata-files, which are named polygons describing exactly what an area-code represents in terms of latitudes and longitudes connected by straight lines. RstoxBase provides the function _RstoxBase::DefineStratumPolygon_ that allows custom spatial strata-system / area codes to be defined. Since FDA frequently make use of standardized areas for reporting, some of these are provided by RstoxFDA as built-in options via the function _RstoxFDA::LoadFdaStratumPolygon_. These area definitions are compatible with those provided by _RstoxBase::DefineStratumPolygon_, and can be used in any function that accepts **RstoxBase::StratumPolygon**, but alleviates the need to obtain resource files for common coding systems such as the statistical areas used by ICES or FDIR.

In general, FDA make use of these area definitions for several purposes besides representing stratification of samples. In particular they are used to impute positions for data where only area-codes are recorded, to annotate data with spatial domain information that is used in reporting, and as post-stratified categorical variables or covariates in estimation models. So in contrast to StoX survey-estimation projects, FDA-projects may well end up reading or loading several area-definitions.

## Imputing positions
Positions can be imputed from area codes, when that gives acceptable precision. That is, a single position within each area (typically the centroid) are used to represent the area and is injected into data that does not have exact positional information. This is most commonly done in order to convert back to other area codes. While that may seem like a roundabout approach, it offers great flexibility in how area codes or stratification systems used in models are defined, and makes it easy to annotate several area-code definitions to data, which is often necessary for purposes of reporting. For these use cases it is important to check that the area codes that positions are imputed from are not too much coarser, and preferably contained within, the area codes that will later be inferred from imputed positions. If area definitions are loaded into StoX via _RstoxFDA::LoadFdaStratumPolygon_ or _RstoxBase::DefineStratumPolygon_, the overlap between different areas may be visualized by selecting the 'Show in Map'-option.

In order to impute positions, the coupling between area codes and positions must first be defined. This is done with the function _RstoxFDA::DefineAreaPosition_. If area codes are defined with polygons of sufficient quality using _RstoxFDA::LoadFdaStratumPolygon_ or _RstoxBase::DefineStratumPolygon_, _RstoxFDA::DefineAreaPosition_ can use these as input and provide a mapping between area codes and positions that is derived directly from the polygons. Otherwise, such a mapping must be provided as a resource file. Missing positions may then be filled into samples (**RstoxData::BioticData**) with _RstoxFDA::SetAreaPositionsBiotic_ or added as new columns to landings (**RstoxData::StoxLandingData**) with _RstoxFDA::AddAreaPositionStoxLanding_. Note that since the format **RstoxData::BioticData** already have columns for positions, _RstoxFDA::SetAreaPositionsBiotic_ fills in missing values in an already existing column, and this happens before converting to **StoxBioticData**. Landings don't have exact positions, so for the case of landings the positions are added to **RstoxData::StoxLandingData**.

## Annotating area codes
For data with positional information, columns with area codes may be added using _RstoxFDA::AddStratumStoxBiotic_ and _RstoxFDA::AddStratumStoxLanding_. The latter provides several options for the name of the column to be added, which can be used to either add a column, or change the existing area column in **RstoxData::StoxLandingData**. By adding several AddStratumStoxLanding-processes to the StoX project, one area-column ('Stratum') can be used as strata or model covariate, and one area-column ('Area') can be used as a domain definition for reporting purposes.

## Neighbour matrix
Reca offers a special kind of random effect, called a CAR-effect, which is often used for spatial covariates and requires a definition of neighborhood between for instance areas. If direct spatial neighbors are desired, and if polygons are provided with sufficient quality, the function _RstoxFDA::DefineCarNeighbours_ can be used to calculate such neighborhood-information from areas defined with _RstoxFDA::LoadFdaStratumPolygon_ or _RstoxBase::DefineStratumPolygon_.

##  Example: StoX-project
Area codes can be used for imputing positions and annotating area codes to data with position, as exemplified by the the sequence of processes in the table below:

| Process name | Function | Description |
|-------|------|---------|--------|
| ReadSamples | ReadBiotic | Reads in **RstoxData::BioticData** with biological samples from NMDbiotic xml-files |
| LoadFdir | LoadFdaStratumPolygon | Loads built-in polygons for the FDIR areas |
| AreaPosition | DefineAreaPosition | Makes a table mapping areas to positions based on polygons defined by LoadFdir |
| SetPositions | SetAreaPositionsBiotic | Set 'latitudestart' and 'longitudestart' on the table 'fishstation' in **RstoxData::BioticData** based on the fields 'area', 'location' and 'system', using the table defined by LoadFdir. Note that this is done before converstion to StoxBiotic for sample data |
| StoxBiotic | StoxBiotic | Converts **RstoxData::BioticData** to **RstoxData::StoxBiotic** |
| ReadLandings | ReadLandingFDA | Reads in **RstoxData::LandingData** with sales-notes data from one of several available file formats |
| StoxLanding | StoxLanding |Converts **RstoxData::LandingData** to the aggregated **RstoxData::StoxLandingData** |
| AddPositions | AddAreaPositionStoxLanding | Add the columns 'Latitude' and 'Longitude' based on the field 'Area', using the table defined by LoadFdir. Note that this is done after conversion to StoxLanding for landing-data |
| DefineStrata | DefineStratumPolygon | Read strata definition from custom polygon file |
| AddStratumSamples | AddStratumStoxBiotic | Add the column 'Stratum' to **RstoxData::StoxBioticData** based on definition read by DefineStrata |
| AddStratumLandings | AddStratumStoxLanding | Add the column 'Stratum' to **RstoxData::StoxLandingData** based on definition read by DefineStrata |
| LoadICES | LoadFdaStratumPolygon | Loads built-in polygons for the ICES areas |
| AddDomainLandings | AddStratumStoxLanding | Change the column 'Area' in **RstoxData::StoxLandingData** based on definition read by LoadICES |

# Harmonising codes
In FDA samples must typically be assigned to partitions of the census. Typically the landings are partitioned by a spatial component (area), gear and a temporal component (period), and these variables must be represented with the same coding systems in both samples (**RstoxData::StoxBioticData**) and landings (**RstoxData::StoxLandingData**). The samples and landings may be provided with different coding systems for instance for gear. Or data may be provided with different resolution, or too high resolution to adequately utilize the available samples. If that is the case, re-coding and re-grouping of codes are necessary, and functions for achieving that is introduced below.

## Spatial component (Area codes)
The section 'Area' describes how the spatial component can be handled with _RstoxFDA::AddStratumStoxBiotic_ and _RstoxFDA::AddStratumStoxLanding_, which result in both samples and landings being annotated with the same area codes. 

## Temporal component (Period codes)
The temporal component can be handled with the functions _RstoxFDA::AddPeriodStoxBiotic_ and _RstoxFDA::AddPeriodStoxLanding_. Similarly to _RstoxFDA::AddStratumStoxLanding_, _RstoxFDA::AddPeriodStoxLanding_ offers several options for the name of the column that the temporal category (period) will be added to, so that it is possible to add one variable ('Period') for use as strata or covariate, and one varaible ('ReportPeriod') for defining domains for reporting. Both _RstoxFDA::AddPeriodStoxBiotic_ and _RstoxFDA::AddPeriodStoxLanding_ requires the definition of the temporal category to be provided by the function _RstoxFDA::DefinePeriod_. They can both refer the same instance of _RstoxFDA::DefinePeriod_, and in fact that is advisable, in order to ensure that the same coding system is used. _RstoxFDA::DefinePeriod_ defines a mapping between dates and periods, and allows for 'Quarter', 'Month' or custom periods to be defined.

## Gear codes
Gear codes can be handled with the functions _RstoxFDA::AddGearGroupStoxBiotic_ and _RstoxFDA::AddGearGroupStoxLanding_, which adds a new column 'GearGroup' to these tables. The mapping between gear codes are provided as a resource file for both of these, and is read into Stox with the function _RstoxData::DefineTranslation_, setting up a mapping between the variable 'Gear' and 'NewValue'. 'NewValue' is the harmonized code that will be used for both samples and landings. _RstoxData::DefineTranslation_ allows several 'Gear' values to be mapped to the same 'NewValue', so that gears may be grouped in this fashion. _RstoxFDA::AddGearGroupStoxBiotic_ and _RstoxFDA::AddGearGroupStoxLanding_ adds the codes from 'NewValue' in a new column named 'GearGroup'. As for the spatial and temporal components, estimation and reporting may be treated differently. This can be achieved by re-coding the 'Gear'-variable of either **RstoxData::StoxBiotic** or **RstoxData::StoxLanding** _after_ 'GearGroup' has been added. The functions _RstoxData::TranslateStoxBiotic_ and _RstoxData::TranslateStoxLanding_ may be utilised for that purpose. 

When the source for the FDA is landings data from FDIR and samples from IMR (NMDbiotic), one cannot immediately use a common gear-definition for both landings and samples. The approach outlined above rather defines common gear groups via two instances of _RstoxData::DefineTranslation_. Another approach is to first translate the 'Gear'-variable for the samples to the same coding system used by landings, and then use a common gear-grouping to add 'GearGroup'. Particularly if 2-digit gear codes are used for the sample-data, one needs to be careful with this approach to make sure that the re-coding of the sample-gears are complete. Otherwise the 'Gear'-column in the sample data will contain two-digit codes from two different coding-standards.

##  Example: StoX-project
Coding systems may be harmonized between samples and landings by a sequence of StoX-processes exemplified in the table below:

| Process name | Function | Description |
|-------|------|---------|--------|
| DefineGearGroupLandings | DefineTranslation | Reads resource file mapping gear to gear-groups for landings |
| DefineGearGroupSamples | DefineTranslation | Reads resource file mapping gear to the same gear-groups  for samples |
| DefinePeriod | DefinePeriod | Defines mapping from date to quarter |
| ReadSamples | ReadBiotic | Reads in **RstoxData::BioticData** with biological samples from NMDbiotic xml-files |
| StoxBiotic | StoxBiotic | Converts **RstoxData::BioticData** to **RstoxData::StoxBioticData** |
| AddGearSamples | AddGearGroupStoxBiotic | Add the column 'GearGroup' to **RstoxData::StoxBioticData** using the definition from DefineGearGroupSamples |
| AddPeriodSamples | AddPeriodStoxLanding | Add the column 'Period' to **RstoxData::StoxBioticData** using the definition from DefinePeriod. |
| ReadLandings | ReadLandingFDA | Reads in **RstoxData::LandingData** with sales-notes data from one of several available file formats |
| StoxLanding | StoxLanding | Converts **RstoxData::LandingData** to the aggregated **RstoxData::StoxLandingData** |
| AddGearLandings | AddGearGroupStoxLanding | Add the column 'GearGroup' to **RstoxData::StoxLandingData** using the defintion from DefineGearGroupLandings |
| AddPeriodLandings | AddPeriodStoxLanding | Add the column 'Period' to **RstoxData::StoxLandingData** using the definition from DefinePeriod. |

# Stock-splitting (coastal cod analysis)
At IMR some species are stock-typed based on otolith readings, and Reca provides an analysis to estimate the catch at age in two domains defined by these stocks. This is an example that relies on variables not supported by **RstoxData::StoxBiotic**, and the column with the stock-types need to be added by using _RstoxData::AddToStoxBiotic_. In addition some model parameters for that analysis can be defined in the StoX-baseline, which is described in the section 'Reca parameters'

# Reca parameters
For estimation with the package Reca, some complex parameters for the statistical models must be defined in the Stox-Baseline, if Reca is configured to use them. These are:

_RstoxFDA::DefineStockSplittingParameters_
 : Defines uncertainty parameters for stock-splitting (coastal-cod analysis).
 
_RstoxFDA::DefineAgeErrorMatrix_
 : Defines uncertainty parameters for age readings.
 
_RstoxFDA::DefineCarNeighbours_
 : Define the neighbour-matrix for the CAR-variable.
 
All of these require resource files, except _RstoxFDA::DefineCarNeighbours_, which may also be calculated from area defintions, as also described in the section 'Area codes'.

# Analysis
The functions introduced in this vignette should enable you to use IMR samples and FDIR landings to obtain **RstoxData::StoxBiotic** and **RstoxData::StoxLanding** data with complete records and harmonized code, ready for further analysis or reporting. Relevant analysis and reporting functions are introduced in other vignettes. See for instance [Stox Fisheries overview (report)](Stox-FDA-overview.html) for obtaining reports on fishing activity and how it is sampled, or [Stox-Reca (analysis)](stox-reca.html) for estimating Catch At Age with Reca.